Verifying Cache Coherence Protocols for an Infinite Family of Cache Hierarchies

### 2014-02-06 paper/RelatedWork.tex
stuff

\section{Related work}
\label{relatedWork}
These are lifted as is from that guy. Need to figure out how to restate them, and make it more concise.

For verification of high level specifications, modern industrial practice
consists of modeling small instances of the protocols, e.g., three CPUs
handling two addresses and one bit of data, in terms of interleaving atomic
steps, in guard/action languages such as Murphi \cite{tla} or TLA+ \cite{tla}, and
exploring the reachable states through explicit state enumeration.
Symbolic methods, e.g., BDD [31] or SAT [129, 142], are yet to succeed for the
verification of cache coherence protocols.
Monolithic formal verification methods – methods that treat the protocol as a whole – have
been used fairly routinely for verifying cache coherence protocols from the early 1990s
[10 (2 nodes, murphi)]. However, these monolithic techniques will not be able to handle the very
large state space of hierarchical protocols. Compositional verification techniques are essential.

>>>>

\section{Related work}
\label{relatedWork}
Formal verification of cache coherence protocols, among other hardware systems,
is mainly done using model checking. Explicit state model checking tools like
Murphi \cite{murphi} or TLC \cite{tlc} have been used in practise only to
handle small and simplified systems, \eg{} a single level cache hierarchy with
three CPUs handling two addresses and one bit of data. Symbolic model checking
techniques have fared slightly better, McMillan \etal{} have verified a 2-level
MSI protocol based on Gigamax distributed multiprocessor using SMV.
Optimizations on these techniques (like partial order reduction \cite{part},
symmetry reduction \cite{sym1, sym2}, compositional reasoning \cite{somethink}, \etc
scale the model checking approach further, but they still haven't been able to
verify realistic hierarchical coherence protocols.

Theorem prover based approaches have also been used to verify cache coherence
protocols. FLASH coherence protocol \cite{flash} has been verified using an
approach called Aggregation of Distributed transactions \cite{park}.  Several
state-transitions in the implementation accessing localized states intuitively
correspond to a single atomic transition accessing all the states globally. In
this approach, the user has to provide a proof of a mapping between the
implemantion transitions to the aggregated atomic transition; once this is
done, a model checker is used to verify the aggregated transition resulting in
verification of the implementation. We believe that the difficult part in
proving a protocol is precisely the proof of such a mapping from an
implementation-level transition to a specification-level transitions accessing
states globally, and their work does not present any approach to go about it.

%These are lifted as is from that guy. Need to figure out how to restate them, and make it more concise.
%
%For verification of high level specifications, modern industrial practice
%consists of modeling small instances of the protocols, e.g., three CPUs
%handling two addresses and one bit of data, in terms of interleaving atomic
%steps, in guard/action languages such as Murphi \cite{tla} or TLA+ \cite{tla}, and
%exploring the reachable states through explicit state enumeration.
%Symbolic methods, e.g., BDD [31] or SAT [129, 142], are yet to succeed for the
%verification of cache coherence protocols.
%Monolithic formal verification methods – methods that treat the protocol as a whole – have
%been used fairly routinely for verifying cache coherence protocols from the early 1990s
%[10 (2 nodes, murphi)]. However, these monolithic techniques will not be able to handle the very
%large state space of hierarchical protocols. Compositional verification techniques are essential. 
%
%For coherence protocols which only consider one level (or nonhierarchical protocols),
%various techniques have been proposed for the verification. These techniques basically
%include explicit state enumeration \cite{sci-cache-coherence} and symbolic model checking \cite{gigamax}
%Opti-
%mizations on these techniques include partial order reduction [24], symmetry reduction
%[25, 72], compositional reasoning [96], predicate abstraction [22, 49, 82], etc. There is
%also a large body of work on parameterized verification [42, 53, 54, 100, 117]. However,
%none of them have shown the capability to verify hierarchical coherence protocols with
%realistic complexity.
%
%McMillan et al. [97, 102] modeled a 2-level MSI coherence protocol based on the
%Gigamax distributed multiprocessor [102]. In the protocol, bus snooping is used in both
%levels. SMV [97] was used to verify this protocol with two clusters each having six
%processors, for both safety and liveness properties. In terms of the complexity, many
%realistic details are abstracted away from the protocol so that the protocol does not have
%the typical complexity that a hierarchical coherence protocol has.
%
